Geodesic Flow on the Necker Cube Surface

### 2017-07-05 paper.tex
070517

\subsection{The Staircase Model}
At first glance, $\tilde{\mathbf{M}}$ does not seem to be a square-tiled surface. However, observe that the surface can be reconstructed by a sequence of cutting and gluing operations and realized in the following way:

\begin{figure}[H]
\centering
\includesvg[width=4in]{mtildastaircase}
\caption{$\tilde{\mathbf{M}}$ as a staircase with vertices and phases identified.}
\end{figure}

Every edge on the staircase model is glued by translation to another by at most a distance of 2. (Each square is unit). More importantly, the staircase has a very natural cylindrical decomposition.

@@

The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions. The modulus of the cylinder, $\mu$, is given by the ratio of the cylinder's circumference and width, $\mu{}={}w/c$. A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by Dehn-twists of magnitude equal to that of the lowest common multiple of the inverse moduli, $\mu^{-1}$, of the horizontal and vertical decompositions.

\begin{figure}[H]
\centering
\includesvg[width=2.6in]{cylinderskew}
\caption{Stabilizing Dehn-twists of the horizontal/vertical cylinders}
\end{figure}

In this case we are lucky and the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$. We can see that any parabolic skews of magnitude 2 preserve the boundaries of the cylinders. Consequently, our surface is stabilized by the following matrices:

\begin{equation}
\left[ \hspace{1mm} \begin{matrix}
				1 & \pm 2\\
				0 & 1
			\end{matrix}\hspace{1mm}\right],
			\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							\pm 2 & 1
						\end{matrix}\hspace{1mm}\right]
\end{equation}

\newpage
\subsection{Other symmetries of $\tilde{\mathbf{M}}$}

@@

\subsection{Computing the Veech Group of $\tilde{\mathbf{M}}$}
\subsection{Slope Normalization Algorithm}
\subsection{Proof of Theorem}

>>>>

\subsection{The Staircase Model}
Observe that $\tilde{\mathbf{M}}$ can be reconstructed by a series of cutting and gluing operations and realized as a square tiled staircase:

\begin{figure}[H]
\centering
\includesvg[width=4in]{mtildastaircase}
\caption{$\tilde{\mathbf{M}}$ as a staircase with vertices $x_i$ and phases identified.}
\end{figure}

Every edge pair on the staircase model has an exact distance of two. (Each square is unit). This gives way to a very natural cylindrical decomposition.

@@

The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions. All families of saddle connections on the surface intersect exactly two others.

\begin{Def}
The \textbf{family of horizontal cylinders} of $\tilde{\mathbf{M}}$, denoted $\Delta_{i}$ where $1\leq i \leq 6$, are cylinders formed by saddle connections made between singularities $((x_3,x_2)\cup(x_2,x_3)$, $(x_3,x_4))$ when i is odd, and $((x_1,x_4)\cup(x_2,x_4)$, $(x_1,x_2))$ when i is even. \newline
Likewise, the \textbf{family of vertical cylinders}, denoted $\varepsilon_j$ $(1\leq j 
\leq 6)$, are formed by connections $((x_2,x_3)$, $(x_2,x_1)\cup(x_1,x_2))$ when j is odd, and  $((x_4,x_1)$, $(x_4,x_3)\cup(x_3,x_4))$ when j is even.
\end{Def}

Horizontal and vertical cylinders of the surface are disjoint within their respective sets, and the union of both sets is the surface $\tilde{\mathbf{M}}$.

  Recall that the modulus of a cylinder, $\mu$, is given by the ratio of the cylinder's width to its circumference, $\mu{}={}w/c$. A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by parabolic matrices with magnitude $\lambda$ equal to that of lowest common multiple of the inverse moduli, $\mu^{-1}$, of both horizontal and vertical decompositions. [cite Veech 89?]

\begin{figure}[H]
\centering
\includesvg[width=2.6in]{cylinderskew}
\caption{Stabilizing Dehn-twists of the horizontal/vertical cylinders}
\label{fig:skew}
\end{figure}

In this case, the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$. We can see that any parabolic skews of magnitude $\lambda=\pm2$, since $\mu^{-1}{}=2$, preserve the boundaries of the cylinders. Figure $\ref{fig:skew}$ showcases how Dehn-twists of this magnitude followed by an Affine transformation matches these boundaries again. Consequently, our surface is stabilized by the matrices

\begin{equation}
\left[ \hspace{1mm} \begin{matrix}
				1 & \pm 2\\
				0 & 1
			\end{matrix}\hspace{1mm}\right],\text{ and }
			\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							\pm 2 & 1
						\end{matrix}\hspace{1mm}\right].
\end{equation}

\newpage
\subsection{Other symmetries of $\tilde{\mathbf{M}}$}
A major benefit of $\tilde{\mathbf{M}}$ being tiled by unit squares is that the surface has all of the same symmetries of the square. The action of $\mathbb{D}_4$ on $\tilde{\mathbf{M}}$ preserves its boundaries up to translation.

\subsection{Computing the Veech Group of $\tilde{\mathbf{M}}$}

@@

\subsection{Slope Normalization Algorithm}
\subsection{Proof of Theorem}
